---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook:
    df_print: paged
---

```{r setup}
library(conditionalsads)
```

This analysis will also source functions from `meteR` and `feasiblesads`. 

## 1. Load data
```{r generate Portal}
setwd(here::here())
portal_plants <- load_portal_plants(download = F)

```

## 2. Generate samples from constraints

For each sample, generate `nsamples` samples from the feasible set and from the METE distribution. 

**These will be the sample_sads for comparison.**


```{r sample constraints}

nsamples <- 100

constraint_samples <- list()

for(i in 1:nrow(toy_fs)) {
  s = length(which(!is.na(toy_fs[i, 3:ncol(toy_fs)])))
  n = sum(toy_fs[i,3:ncol(toy_fs)], na.rm = T)
  this_fs <- sample_feasibleset(s = s, n = n, nsamples)
  this_mete <- sample_METE(s = s, n= n, nsamples)
  
  these_constraint_samples <- list(this_fs, this_mete)
  
  constraint_samples[[i]] <- these_constraint_samples
  
  rm(this_fs)
  rm(this_mete)
  rm(these_constraint_samples)
}


```


## 3. Calculate test statistics 

* Empirical R2 and sample R2 compared to most-likely constraint vector.
* Empirical and sample evenness (Evar and Simpson's) and skew.

```{r R2}

fs_r2 <- list()
mete_r2 <- list()

fs_evar <- list()
fs_simp <- list()
fs_skew <- list()

mete_evar <- list()
mete_simp <- list()
mete_skew <- list()

for(i in 1:nrow(toy_fs)) {
  s = length(which(!is.na(toy_fs[i, 3:ncol(toy_fs)])))
  n = sum(toy_fs[i,3:ncol(toy_fs)], na.rm = T)
  fs_r2[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                              sampled_rads = constraint_samples[[i]][[1]],
                              stat = "r2", constraint = get_fs_ct(s, n, nsamples = nsamples))
  mete_r2[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "r2", constraint = get_mete_prediction(s, n))

  fs_evar[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "evar")
  fs_simp[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "simpson")
  fs_skew[[i]] <-  get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "rad_skew")
  
  mete_evar[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[2]],
                                stat = "evar")
  mete_simp[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[2]],
                                stat = "simpson")
  mete_skew[[i]] <- get_stat_list(empirical_rad = as.integer(toy_fs[i, 3:(s+2)]),
                                sampled_rads = constraint_samples[[i]][[2]],
                                stat = "rad_skew")
  
  
  rm(s)
  rm(n)
}

```


## 4. Get quantiles of empirical test statistics vs. constraints

```{r get quantiles}
fs_r2_quantile <- vapply(fs_r2, FUN = test_quantile, FUN.VALUE = 1)
fs_evar_quantile <- vapply(fs_evar, FUN = test_quantile, FUN.VALUE = 1)
fs_simp_quantile <- vapply(fs_simp, FUN = test_quantile, FUN.VALUE = 1)
fs_skew_quantile <- vapply(fs_skew, FUN = test_quantile, FUN.VALUE = 1)


mete_r2_quantile <- vapply(mete_r2, FUN = test_quantile, FUN.VALUE = 1)
mete_evar_quantile  <- vapply(mete_evar, FUN = test_quantile, FUN.VALUE = 1)
mete_simp_quantile  <- vapply(mete_simp, FUN = test_quantile, FUN.VALUE = 1)
mete_skew_quantile  <- vapply(mete_skew, FUN = test_quantile, FUN.VALUE = 1)


toy_fs <- cbind(toy_fs, fs_r2_quantile, fs_evar_quantile, fs_simp_quantile, fs_skew_quantile, mete_r2_quantile, mete_evar_quantile, mete_simp_quantile, mete_skew_quantile)

if(FALSE) {
  write.csv(toy_fs, "toy_fs_done.csv")
}

```


## 5. Compare quantiles across experimental treatments

```{r fs plots, echo = F}

library(ggplot2)


for(i in 1:4) {

  this_h <- levels(toy_fs$h)[i]
  
  this_data <- toy_fs %>%
    dplyr::filter(h == this_h)
  
  r2_plot <-  ggplot(data = this_data, aes(x = trtmnt, y = fs_r2_quantile)) +
    geom_jitter(aes(x = this_data$trtmnt, y = this_data$fs_r2_quantile)) +
    ggtitle(paste0(this_h, " r2")) +
    theme_bw()
  
  evar_plot <-  ggplot(data = this_data, aes(x = trtmnt, y = fs_evar_quantile)) +
    geom_jitter(aes(x = this_data$trtmnt, y = this_data$fs_evar_quantile)) +
    ggtitle(paste0(this_h, " evar")) +
    theme_bw()
  
  simp_plot <-  ggplot(data = this_data, aes(x = trtmnt, y = fs_simp_quantile)) +
    geom_jitter(aes(x = this_data$trtmnt, y = this_data$fs_simp_quantile)) +
    ggtitle(paste0(this_h, " Simpson evenness")) +
    theme_bw()
  
  skew_plot <-  ggplot(data = this_data, aes(x = trtmnt, y = fs_skew_quantile)) +
    geom_jitter(aes(x = this_data$trtmnt, y = this_data$fs_skew_quantile)) +
    ggtitle(paste0(this_h, " skewness")) +
    theme_bw()
  
print(gridExtra::grid.arrange(r2_plot, evar_plot, simp_plot, skew_plot,
                          nrow = 1))
}


```

```{r mete plots}

```
