---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook:
    df_print: paged 
---

```{r setup}
library(conditionalsads)
```

This analysis will also source functions from `meteR` and `feasiblesads`. 

## 1. Load data
```{r generate Portal, eval =F}
setwd(here::here())
portal_plants <- process_portal_plants(load_portal_plants(download = F))

rm_list = ls()
rm(list = rm_list[ which(rm_list != "portal_plants")])
rm(rm_list)

plant_abund <- portal_plants[[2]]
plant_abund <- as.matrix(plant_abund)

```
```{r load plant data, echo = F}
setwd(here::here())
load('analysis/report/portal_plants.Rds')
```

## 2. Generate samples from constraints

For each sample, generate `nsamples` samples from the feasible set and from the METE distribution. 

**These will be the sample_sads for comparison.**


```{r sample constraints, eval = F}

nsamples <- 2

constraint_samples <- list()

for(i in 1:nrow(plant_abund)) {
  s = length(which(!is.na(plant_abund[i, ])))
  n = sum(plant_abund[i,], na.rm = T)
  this_fs <- sample_feasibleset(s = s, n = n, nsamples)
  this_mete <- sample_METE(s = s, n= n, nsamples)
  
  these_constraint_samples <- list(this_fs, this_mete)
  
  constraint_samples[[i]] <- these_constraint_samples
  print(i)
  rm(this_fs)
  rm(this_mete)
  rm(these_constraint_samples)
  save.image('sampling_constraints.RData')
}


```


## 3. Calculate test statistics 

* Empirical R2 and sample R2 compared to most-likely constraint vector.
* Empirical and sample evenness (Evar and Simpson's) and skew.

```{r R2, eval =F}

for(i in 1:nrow(plant_abund)) {
  
 this_rad = as.integer(plant_abund[i, ])
  this_rad = na.omit(this_rad)
  s = length(this_rad)
  n = sum(this_rad)
  
  fs_constraint = get_fs_ct(s, n, nsamples = nsamples, newsamples = F, oldsamples = constraint_samples[[i]][[1]])
  mete_constraint  = get_mete_prediction(s, n)
  
  fs_r2[[i]] <- get_stat_list(empirical_rad = this_rad,
                              sampled_rads = constraint_samples[[i]][[1]],
                              stat = "r2", constraint = fs_constraint)
  fs_kl[[i]] <- get_stat_list(empirical_rad = this_rad,
                              sampled_rads = constraint_samples[[i]][[1]],
                              stat = "kl_div", constraint = fs_constraint)
  
  
  mete_r2[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "r2", constraint = mete_constraint)
  
  
  mete_kl[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "kl_div", constraint = mete_constraint)

  fs_evar[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "evar")
  fs_simp[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "simpson")
  fs_skew[[i]] <-  get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[1]],
                                stat = "rad_skew")
  
  mete_evar[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[2]],
                                stat = "evar")
  mete_simp[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[2]],
                                stat = "simpson")
  mete_skew[[i]] <- get_stat_list(empirical_rad = this_rad,
                                sampled_rads = constraint_samples[[i]][[2]],
                                stat = "rad_skew")
  
  poilogs[[i]] <- rad_poilog_cs(this_rad)
  print(i)
  save.image('getting_stats.RData')
  
  rm(s)
  rm(n)
}


```


## 4. Get quantiles of empirical test statistics vs. constraints

```{r get quantiles, eval = F}
fs_r2_quantile <- vapply(fs_r2, FUN = test_quantile, FUN.VALUE = 1)
fs_kl_quantile <- vapply(fs_kl, FUN = test_quantile, FUN.VALUE =1)
fs_evar_quantile <- vapply(fs_evar, FUN = test_quantile, FUN.VALUE = 1)
fs_simp_quantile <- vapply(fs_simp, FUN = test_quantile, FUN.VALUE = 1)
fs_skew_quantile <- vapply(fs_skew, FUN = test_quantile, FUN.VALUE = 1)


mete_r2_quantile <- vapply(mete_r2, FUN = test_quantile, FUN.VALUE = 1)
mete_kl_quantile <- vapply(mete_kl, FUN = test_quantile, FUN.VALUE =1)
mete_evar_quantile  <- vapply(mete_evar, FUN = test_quantile, FUN.VALUE = 1)
mete_simp_quantile  <- vapply(mete_simp, FUN = test_quantile, FUN.VALUE = 1)
mete_skew_quantile  <- vapply(mete_skew, FUN = test_quantile, FUN.VALUE = 1)


# Poilog pars

pull_poilog <- function(pl_list, item) {
  return(pl_list[[item]])
}

poilog_expmu <- vapply(poilogs, FUN = pull_poilog, FUN.VALUE = 1, item = 1)
poilog_sig <- vapply(poilogs, FUN = pull_poilog, FUN.VALUE = 1, item = 2)

plant_abund_results <- cbind(portal_plants[[1]], plant_abund, fs_r2_quantile,fs_kl_quantile, fs_evar_quantile, fs_simp_quantile, fs_skew_quantile, mete_r2_quantile, mete_kl_quantile, mete_evar_quantile, mete_simp_quantile, mete_skew_quantile, poilog_expmu,
                poilog_sig)

#if(FALSE) {
  write.csv(plant_abund_results, "plants_2samples_done.csv")
  save.image('plants_2samples.RData')
#}

```


## 5. Compare quantiles across experimental treatments

```{r fs plots, echo = F, fig.width = 4, fig.height=20}
# load('plants_done.RData')
library(ggplot2)
setwd(here::here())
plant_abund_results <- read.csv('plants_mete.csv')
for(i in 1:2) {

  this_season <- unique(plant_abund_results$season)[i]
  
  this_data <- plant_abund_results %>%
    dplyr::filter(season == this_season)
  
  r2_plot <-  ggplot(data = this_data, aes(x = trmt, y = fs_r2_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$fs_r2_quantile)) +
    ggtitle(paste0(this_season, " r2 - FS")) +
    theme_bw()
  
  kl_plot <-  ggplot(data = this_data, aes(x = trmt, y = fs_kl_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$fs_kl_quantile)) +
    ggtitle(paste0(this_season, " kl - FS")) +
    theme_bw()
  
  evar_plot <-  ggplot(data = this_data, aes(x = trmt, y = fs_evar_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$fs_evar_quantile)) +
    ggtitle(paste0(this_season, " evar - FS")) +
    theme_bw()
  
  simp_plot <-  ggplot(data = this_data, aes(x = trmt, y = fs_simp_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$fs_simp_quantile)) +
    ggtitle(paste0(this_season, " Simpson evenness - FS")) +
    theme_bw()
  
  skew_plot <-  ggplot(data = this_data, aes(x = trmt, y = fs_skew_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$fs_skew_quantile)) +
    ggtitle(paste0(this_season, " skewness - FS")) +
    theme_bw()
  
    r2_mete_plot <-  ggplot(data = this_data, aes(x = trmt, y = mete_r2_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$mete_r2_quantile)) +
    ggtitle(paste0(this_season, " r2 - METE")) +
    theme_bw()
  
  kl_mete_plot <-  ggplot(data = this_data, aes(x = trmt, y = mete_kl_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$mete_kl_quantile)) +
    ggtitle(paste0(this_season, " kl - METE")) +
    theme_bw()
  
  evar_mete_plot <-  ggplot(data = this_data, aes(x = trmt, y = mete_evar_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$mete_evar_quantile)) +
    ggtitle(paste0(this_season, " evar - METE")) +
    theme_bw()
  
  simp_mete_plot <-  ggplot(data = this_data, aes(x = trmt, y = mete_simp_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$mete_simp_quantile)) +
    ggtitle(paste0(this_season, " Simpson evenness - METE")) +
    theme_bw()
  
  skew_mete_plot <-  ggplot(data = this_data, aes(x = trmt, y = mete_skew_quantile)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$mete_skew_quantile)) +
    ggtitle(paste0(this_season, " skewness - METE")) +
    theme_bw()
    poilog_mu_plot <- ggplot(data = this_data, aes(x = trmt, y = poilog_expmu)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$poilog_expmu)) +
    ggtitle(paste0(this_season, " poilog_expmu")) +
    theme_bw()
  
   poilog_sig_plot <- ggplot(data = this_data, aes(x = trmt, y = poilog_sig)) +
    geom_jitter(aes(x = this_data$trmt, y = this_data$poilog_sig)) +
    ggtitle(paste0(this_season, " poilog_sig")) +
    theme_bw()
  
print(gridExtra::grid.arrange(r2_plot, r2_mete_plot, kl_plot, kl_mete_plot,
                              evar_plot, evar_mete_plot, simp_plot,  simp_mete_plot, skew_plot,
                               skew_mete_plot,
                          nrow = 5))
}


```
